Brownian motion of main-belt asteroids on human timescales

### 2025-02-23 main.tex
Completed draft with abstract.

\begin{abstract}
  xxx
\end{abstract}

@@

High-precision measurements of positions of Solar System bodies have
great value to multiple scientific pursuits.  They can reveal
gravitational accelerations due to undiscovered bodies such as a
hypothetical ``Planet X'' \citep[\eg][]{holmanP9,inpopP9,trojans,occultations} or passing primordial
black holes \citep{pbh}, determine
the masses of known bodies \citep[\eg][]{goffin,baer}, test the laws of physics, and characterize 
the non-gravitational forces that drive many forms of planetary
migration, \eg\ the maintainence of the near-Earth asteroid (NEA)
population \citep{yorp}.  While most applications of precision orbit determination
have used the 8 major planets as test particles, there are now $>10^6$
known minor planets available for this task.  These small bodies' astrometric
information content is growing extremely rapidly through large-scale
survey projects that can produce
milliarcsecond-scale accuracy for magnitude-limited small-body populations, such as Gaia \citep{gaiass3},
PanSTARRS,\footnote{\url{https://www2.ifa.hawaii.edu/research/Pan-STARRS.shtml}}
and particularly the upcoming Rubin observatory
\textit{Legacy Survey of Space and Time} (LSST).\footnote{\url{https://rubinobservatory.org/}}  The great majority
of currently-tracked objects are main-belt asteroids (MBAs), so we
wish to investigate the potential power of MBAs as gravitational test
bodies.

The most advanced ephemerides for the Solar System \citep{de440,inpop,pitjeva} consider the
gravitational forces emanating from the Sun, from the major planets and their large
moons, from up to $\approx300$ of the largest individual MBAs, and
from distributed mass annuli representing the smaller members of
 the asteroid belt(s) and Kuiper belt.
Here we ask: \textit{what is the RMS ephemeris error in angular or radial
position a typical MBA accrues because of its gravitational
encounters with other individual MBAs?}  We are interested not in the
perturbations that can be well described by a multipolar ring model
for the full population; rather the Brownian motion that accumulates
from closer encounters with: (a) MBAs that \emph{are not included} in the
ephemeris model at all; and (b) asteroids that \emph{are} included
in an ephemeris model, but have some uncertainty $\sigma_M$ in their masses.
Any such departures became a source of noise in any
inferences that we make from use of these MBAs as
dynamical tracer particles.  We aim to quantify this noise, as a
function of the RMS uncertainty $\sigma_M$ on the masses of individual
asteroids, and assuming that deflectors with $M<\sigma_M$ are not
included in the ephemeris at all.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.6\textwidth]{result.pdf}
  \caption{Each line shows the forecast of the variance of the
    azimuthal position of a Middle Main Belt asteroid caused by discrete
    encounters with absolute magnitude $H$ greater than the value on
    the (lower) $x$ axis, \ie\ the cumulative noise from all MBAs
    fainter (smaller) than the plotted $H_{\rm min.}$  The upper axis gives the $M_{\rm max}$ equivalent to $H_{\rm min}.$
.  The left axis gives values in meters, the right
    the apparent shift in milliarseconds for a heliocentric observer.
    Each line is for a different assumption about which asteroids' deflections are
    part of the ephemeris model instead of being considered as noise: the $\sigma_M$
    value gives the RMS mass uncertainty on modelled MBAs, and the $N$
    is the number of MBAs with $M>\sigma_M$ that could be usefully
    modelled.  The black dashed curve is a conservative estimate of what can be achieved using LSST data to constrain larger MBA's masses.}
  \label{fig:result}
\end{figure}

@@

\subsection{Implications for future investigations}

An important question is: will the positional noise due to unmodelled MBA encounters degrade inferences being made with measurements of tracer MBAs?  If the Brownian noise is below other sources of measurement noise, then the answer must be ``no.''
The per-epoch uncertainties on solar-system object astrometry in Gaia DR3\footnote{We take one Gaia transit as an ``epoch'' here.} are reported to be as low as 0.3~mas for very bright sources ($G=13$), rising to $\approx10$~mas at its $G=21$ limit where most detections lie \citep[][Figure 6]{gaiass3}.  The overall RMS residual to the orbit fits are 5~mas (this is in the ``along-track'' direction).  Thus the Brownian motion of $\approx0.4$~mas we estimate for $\phi$ (over a 10-year period) would be important for the single-epoch error budget of only the brightest Gaia MBAs.  
The Brownian motion is, however, highly correlated across epochs, whereas each epoch's measurement errors are independent.  The Brownian motion noise hence does not average down as the square root of the number of epochs, which is $\approx20$ over 3 years for a typical MBA in the Gaia DR3 catalog.  The Brownian motion might become relevant for analysis of a significant fraction of Gaia MBAs during analysis of the complete 11-year data.  But with LSST estimates of deflector masses in hand c.\ 2035, with $\sigma_\phi\approx0.25$~mas, the (retrospective) use of Gaia data will probably again will have Brownian motion subdominant to measurement errors the fainter few magnitudes of its catalog.

For LSST, the measurement errors on MBA positions will have a floor of 1--2~mas per epoch set by refraction from atmospheric turbulence \citep{willow,trojans}.  Objects with apparent magnitude $m\gtrsim20$ will have larger per-epoch errors due to photon noise in the images.  There will, however, be hundreds of observations taken of each MBA over 10~yr, meaning that Brownian motion might be a significant contributor to inferences from the dynamics of most of its tracer MBAs, if our $\sigma_M$ levels were fixed near current knowledge.
On the other hand, analyses of LSST astrometry of MBAs should certainly take advantage of its ability to improve knowledge of the masses of individual larger MBAs, and lower the ephemeris errors to $\approx0.25$~mas, well below single-epoch atmospheric turbulence noise..  Considering the coherence of the Brownian noise over $O(100)$ LSST observations, the Brownian noise could become comparable to measurement noise for inferences using the brighter tracer MBAs in LSST data, but will remain unimportant for the great majority of LSST's measured MBAs.

An even more ambitious proposal by \citet{occultations} is to track $\approx10^6$ MBAs down to diameters of $\approx1$~km by watching them occult Gaia stars.  The timing of each occultation can locate the center of mass of the target MBA along its direction of motion, with most of the information available from objects yielding uncertainties of 50--100~m on this position.  The Brownian motion noise over a 10-year baseline will in this case become comparable to the per-epoch measurement errors.  Reducing the typical mass uncertainty of individual asteroids to $\sigma_M\approx 10^{-15.4} M_\odot$ would be needed ($N=3000$) to reduce the Brownian noise to negligible levels for this survey.

@@

The extent to which the Brownian motion can confound an inference will depend on how similar the spatio-temporal dependence of the Brownian motion is (as quantified by its covariance matrix) to the signal under test.  If, for example, one is searching for the tidal force of a Planet X, this will be manifested primarily as a collective apsidal precession and a quadrupolar deviation of the orbit from its ellipse that does not evolve with time.  These are quite distinct from the random walk in $\Delta\tau$ that is the dominant effect of Brownian motion.  The Brownian noise will be somewhat orthogonal to the signal under study in the data space, and therefore cause less degradation to the Planet X inference than the per-epoch $\sigma_\phi$ would suggest.  Another example would be a gravitational perturbation whose signature is primarily a nodal precession---we have seen that Brownian motion out of the orbital plane is $>40\times$ smaller than the $\phi$ component, and a nodal precession's observable effect is primarily in this vertical component.

\subsection{Comparison to radiation forces}
As we consider smaller MBAs as test particles, radiation pressure becomes larger relative to gravitational forces.  It is interesting to compare the effects of radiation pressure to those of Brownian motion to see when the former are a bigger source of noise than the latter, or contrarily to see whether Brownian motion interferes with attempts to measure radiation forces on MBAs.  At first order, radiation pressure from the Sun and the reaction from reflected sunlight cause a constant radial force on the MBA.  The ratio of this force to the gravitational force for an asteroid of radius $R,$ geometric albedo $q$, and density $\rho$ is 
\begin{equation}
 \Psi \equiv \frac{F_{\rm rad}}{F_{\rm grav}} = \frac{ (1+q) \pi R^2 L_\odot / 4\pi c}{4\pi GM_\odot R^3\rho/3}
   = \frac{3(1+q)}{16\pi} \frac{L_\odot}{GM_\odot c R \rho} \approx 3\times10^{-10} 10^{0.2(H-16)},
 \end{equation}
 where the last expression adopts our standard values $q=0.25, \rho=2500\,{\rm kg}\, {\rm m}^{-3}.$ The observational consequence of this constant radial force is an $H$-dependent shift $\Delta r = -\Psi a$ at a fixed period of rotation, which does not resemble any expected effect of Brownian motion.  More important is the Yarkovsky effect, in which anisotropic re-emission of the absorbed radiation leads to a net force that is fraction $g_Y$ of the incident radiation pressure.  Taking this force to be in the prograde azimuthal direction, and using our system of units with $a=1$ and $v_c=1,$ this leads to a continuous power $dE/dt=g_Y\Psi$ which in turn implies $da/dt=2g_Y\Psi.$   The mean motion rate is hence $1-3 g_Y\Psi t$ and the accumulated shift in mean anomaly over time $T$ amounts to $\Delta\phi = -3 g_Y \Psi  T^2/2.$  The actual value of the azimuthal Yarkovsky force (torque) depends on the spin axis, the spin rate, and the radiant and thermal properties of the surface, so neither the sign nor amplitude are known for any individual MBA.   Indeed the Yarkovsky effect has not, to our knowledge, been detected on any MBA, but measurements on NEAs suggest that values of $g_Y\lessim0.1$ might be appropriate.  In this case, a rough comparison of the accumulated $\Delta\phi$ from the Yarkovsky effect to the $\sigma_\phi$ of noise from unmodelled Brownian motion, after a time interval $T$, is
 \begin{equation}
   \frac{\Delta\phi}{\sigma_\phi} \approx \frac{ 3 g_Y \Psi T^2/2}{\sqrt{3 T^3 \nIphi}} \approx \sqrt{T/10\,{\rm yr}} 10^{^{0.2(H-20.2)}
     \label{eq:yarkovsky}
   \end{equation}
   if we assume $\nIphi$ from the current knowledge of asteroid masses.   There is thus a break point near $H=20$, or $\approx100$~m diameter: for MBAs larger than this, the ephemeris uncertainties from the Yarkovsky effect will be smaller than those from mutual gravitational encounters. This includes all MBAs detected by Gaia and nearly all detectable by LSST.  Below this size, the Yarkovsky displacement grows larger in a decade and has a better chance of being detectable in the presence of Brownian motion.  Distinguishing the steady quadratic growth of the Yarkovsky $\Delta\phi$ from the random-walk growth of Brownian motion to the same final size requires observational S/N and time resolution significantly better than is needed for detecting either one individually.

@@

In conclusion, the Brownian motion from unmodelled or mis-modelled mutual impulses between MBAs causes RMS azimuthal displacements at the 1~km scale over 10 years, with current knowledge of the masses of large asteroids.  This is large enough to become a significant part of the error budget for many inferences from the larger/brighter asteroids being used as tracer particles from Gaia or LSST images.  If, however, we can use mutual events from LSST to constrain the masses of the $\approx500$ largest  asteroids with $S/N\gtrsim1,$ then the unmodelled Brownian motion shrinks to $\approx250$~m (or 0.1~mas) for MBAs, which becomes insignificantly small for most uses of Gaia or LSST astrometry.  Occultation-based astrometry might, at this stage, have measurement errors that are similar to the uncertainties from Brownian motion on decade time scales.

>>>>

\begin{abstract}
  The explosion of high-precision astrometric data on main-belt asteroids (MBAs) opens the door to their use for many possible inferences on gravitational and non-gravitational forces present in this region.  We estimate the size of positional shifts on MBAs caused by mutual gravitational encounters with other MBAs that are either omitted from ephemeris models or have uncertain mass estimates. In other words, what is the typical Brownian motion among MBAs that cannot be predicted from the ephemeris, and therefore serves as noise on inferences from the MBAs?  We calculate the RMS azimuthal shift $\sigma_\phi$ of this ``Brownian noise''  by numerical estimation of the distribution of impulse sizes and directions among known MBAs, combined with analytical propagation into future positional uncertainties.  The result depends critically on how well the masses of the larger MBAs are known.  At current level of asteroid-mass knowledge, $\sigma_\phi$ rises to $\approx2$~km or $\approx1$~mas over $T=10$~yr, increasing as $T^{3/2}.$  This is large enough to degrade many inferences from Gaia and LSST MBA data, although the LSST data will also improve MBA mass knowledge enough to lower this Brownian uncertainty by $\ge4\times.$  Radial and vertical Brownian noise  at $T=10$~yr are suppressed by factors $\approx7$ and $\approx45,$ respectively, compared to the azimuthal signal after 10~yr, and grow more slowly with time.  We suggest that for full exploitation of Gaia and LSST MBA data, ephemeris models should include the $\approx1000$~largest asteroids as active bodies with free masses, even if not all are well constrained.  This will correctly propagate the uncertainties from mutual encounters into desired inferences, with the RMS value of positional errors from encounters with less-massive MBAs being limited to $\sigma_\phi\approx60$~m or 30~$\mu$as, small enough to ignore until occultation-based position data become ubiquitously available for MBAs.
\end{abstract}

@@

High-precision measurements of positions of Solar System bodies have
great value to multiple scientific pursuits.  They can reveal
gravitational accelerations due to undiscovered bodies such as a
hypothetical ``Planet X'' \citep[\eg][]{holmanP9,inpopP9,trojans,occultations} or passing primordial
black holes \citep{pbh}, determine
the masses of known bodies \citep[\eg][]{goffin,baer}, test the laws of physics, and characterize 
the non-gravitational forces that drive many forms of planetary
migration, \eg\ the maintainence of the near-Earth asteroid (NEA)
population \citep[reviewed by][]{yarkovsky}.  While most applications of precision orbit determination
have used the 8 major planets as test particles, there are now $>10^6$
known minor planets available for this task.  These small bodies' astrometric
information content is growing extremely rapidly through large-scale
survey projects that can produce
milliarcsecond-scale accuracy for magnitude-limited small-body populations, such as Gaia \citep{gaiass3,gaiafpr},
PanSTARRS,\footnote{\url{https://www2.ifa.hawaii.edu/research/Pan-STARRS.shtml}}
and particularly the upcoming Rubin observatory
\textit{Legacy Survey of Space and Time} (LSST).\footnote{\url{https://rubinobservatory.org/}}  The great majority
of currently-tracked objects are main-belt asteroids (MBAs), so we
wish to investigate the potential power of MBAs as gravitational test
bodies.

The most advanced ephemerides for the Solar System \citep{de440,inpop,pitjeva} consider the
gravitational forces emanating from the Sun, from the major planets and their large
moons, from up to $\approx300$ of the largest individual MBAs (typically those that can influence the exquisitely measured ranging to Mars), and
from distributed mass annuli representing the smaller members of
 the asteroid belt(s) and Kuiper belt.
Here we ask: \textit{what is the RMS ephemeris error in angular or radial
position a typical MBA accrues because of its gravitational
encounters with other individual MBAs?}  We are interested not in the
perturbations that can be well described by a multipolar ring model
for the full population; rather the Brownian motion that accumulates
from closer encounters with: (a) MBAs that \emph{are not included} in the
ephemeris model at all; and (b) asteroids that \emph{are} included
in an ephemeris model, but have some uncertainty $\sigma_M$ in their masses.
Any such departures became a source of noise in any
inferences that we make from use of these MBAs as
dynamical tracer particles.  We aim to quantify this noise, as a
function of the RMS uncertainty $\sigma_M$ on the masses of individual
asteroids, and assuming that deflectors with $M<\sigma_M$ are not
included in the ephemeris at all.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.6\textwidth]{result.pdf}
  \caption{Each line shows the forecast of the standard deviation of the
    azimuthal position of a Middle Main Belt asteroid caused by discrete
    encounters over 10~years with absolute magnitude $H$ greater than the value on
    the (lower) $x$ axis, \ie\ the cumulative noise from all MBAs
    fainter (smaller) than the plotted $H_{\rm min.}$  The upper axis gives the $M_{\rm max}$ equivalent to $H_{\rm min}.$
.  The left axis gives RMS values in meters, the right
    the apparent shift in milliarseconds for a heliocentric observer.
    Each line is for a different assumption about the RMS uncertainty $\sigma_M$ in the masses of individual asteroids gained from high-$S/N$ observations of mutual encounters.
Each star marks the dividing point between the $N$ MBAs with $M>\sigma_M,$ and the smaller bodies with no useful measure of their mass ($M<\sigma_M$). The light dotted lines mark the level of azimuthal noise expected from the bodies beyond the most massive $N,$ \ie\ the RMS that an ephemeris model including $N$ bodies would not properly model. 
    The black dashed curve is a conservative estimate of what can be achieved using LSST data to constrain larger MBA's masses. Radial and vertical RMS positions are factors of $\approx 7$ and 45 lower than the azimuthal noise, respectively.}
  \label{fig:result}
\end{figure}

@@

\section{Consequences}
\subsection{Implications for future inferences}

An important question is: will the positional noise due to unmodelled MBA encounters degrade inferences being made with measurements of tracer MBAs?  If the Brownian noise is below other sources of measurement noise, then the answer must be ``no.''
The per-epoch uncertainties on solar-system object astrometry from Gaia\footnote{We take one Gaia transit as an ``epoch'' here.} are reported to be as low as 0.3~mas for very bright sources ($G=13$), rising to $\approx10$~mas at its $G=21$ limit where most detections lie \citep[][Figure 6]{gaiass3}.  The overall RMS residual to the orbit fits are 5~mas (this is in the ``along-track'' direction).  Thus the Brownian motion of $\approx1$~mas we estimate for $\phi$ (over a 10-year period) with current asteroid-mass knowledge would be important for the single-epoch error budget of the brighter Gaia MBAs.  With our forecasted LSST-era knowledge of asteroid masses ($\sigma_M\approx10^{-13.6},$ RMS azimuthal errors $\approx0.25$~mas), the situation will be improved in that only the brightest few MBAs will have per-epoch Brownian motion noise comparable to the per-epoch measurement noise.

For LSST, the measurement errors on MBA positions will have a floor of 1--2~mas per epoch set by refraction from atmospheric turbulence \citep{willow,trojans}.  Objects with apparent magnitude $m\gtrsim20$ will have larger per-epoch errors due to photon noise in the images.  Assuming that LSST inferences will be able to exploit LSST's own determinations of MBA masses, we see that on a per-epoch basis, the Brownian noise will always be subdominant to measurement errors.

An ambitious proposal by \citet{occultations} is to track $\approx10^6$ MBAs down to diameters of $\approx1$~km by watching them occult Gaia stars.  The timing of each occultation can locate the center of mass of the target MBA along its direction of motion, with most of the information available from objects yielding uncertainties of 50--100~m on this position.  The Brownian noise (with LSST-level asteroid knowledge) over a 10-year baseline will in this case be comparable to the per-epoch measurement errors for most observations.  Reducing the typical mass uncertainty of individual asteroids to $\sigma_M\approx 10^{-15.4} M_\odot$ would be needed ($N=3000$) to reduce the Brownian noise to negligible levels for this survey.

These per-epoch comparisons are, however, misleading. The Brownian noise is highly correlated between different epochs of observation of a given asteroid, while the measurement errors are generally uncorrelated, and therefore the impact on inferences is not described solely by comparison of RMS values of each.  The covariance structure of Brownian noise can work either to the benefit or the detriment of the inference.  In an unfavorable case, the eigenvectors of this covariance matrix with the largest uncertainties look similar to the signal of interest in the inference, \ie\ the derivative of the MBA's position with respect to a parameter.  If there is a single dominant eigenvector of the Brownian noise which looks just like our signal, then the effect per-epoch Brownian noise does not average down at all as we obtain $M$ observations, but the measurement noise drops as $\sqrt{M}.$  In a favorable case, all of the strong eigenvectors of the Brownian noise covariance are orthogonal to the signals of interest, and the Brownian noise causes no degradation of the inference, \ie\ its spatiotemporal structure cannot mimic that of the signal we seek.

Consider first the worst-case scenario, where the Brownian noise does not average down at all across observations, but measurement errors gain a factor $\sqrt{M}.$   \citet{gaiafpr} report typically $\sim60$ epochs per asteroid in 66 months, suggesting $O(100)$ epochs per asteroid in the full 11-year observing period.  At Gaia's faint end, this means that Brownian noise with present asteroid knowledge would have comparable impact to the measurement errors, and dominate inferences at the bright end.  LSST will obtain several hundred epochs per asteroid.  With an effective factor 10--20 reduction in the total measurement noise, and assuming $\sigma_\phi\approx0.25$~mas Brownian noise, the latter would become important for the brighter tracer MBAs in LSST data, but inferences with the millions of fainter objects discovered by LSST would still be limited by photon noise.  The occultation program proposed by \citet{occultations} would obtain $<10$ measurements per decade for most MBAs and hence see less degradation of Brownian relative to measurement noise.

At the other extreme are cases where the dominant modes of Brownian noise are orthogonal to the effects of parameters of interest.  
If, for example, one is searching for the tidal force of a Planet X, this will be manifested primarily as a collective apsidal precession and a quadrupolar deviation of the orbit from its ellipse that does not evolve with time.  These are quite distinct from the random walk in $\Delta\tau$ that is the dominant effect of Brownian motion, and therefore Brownian noise with cause less degradation to the Planet X inference than the per-epoch $\sigma_\phi$ would suggest.  Another example would be a gravitational perturbation whose signature is primarily a nodal precession, which is a signal manifesting primarily as out-of-plane motions---we have seen that Brownian motion out of the orbital plane is $>40\times$ smaller than the $\phi$ component and grows at $T{-1/2}$ instead of $T^{3/2},$ so the per-epoch $\sigma_\phi$ of Brownian noise will grossly overestimate its effect on the inference.

@@

A better approach, however, would probably be to increase the number $N$ of MBAs considered as active bodies in the ephemeris fitting, leaving their masses as free parameters.  Although the extreme case of having $N=O(10^6)$ active bodies would be a substantial computational challenge, having $N=10^3$--$10^4$ is not, especially since the MBA-to-MBA perturbations can be adequately calculated using Newtonian dynamics instead of a fully relativistic treatment.  In this approach, all of the covariances between different observations of the same tracer, and between tracers, will be captured in the analysis for deflections due to the $N$ deflectors.  This is true even if the ephemeris fit does not yield $S/N>1$ on every deflector.  If we \emph{choose $N$ to be large enough that the Brownian noise from the unmodelled MBAs is negligible to the error budget,} then we need not explicitly propagate the Brownian noise.

Figure~\ref{fig:result} shows us what $N$ would be required: the noise due to unmodelled deflectors is given by the point where each curve for a given $N$ departs from the upper-right envelope of the curves.   Increasing the number of MBAs in the ephemeris from $N=100\rightarrow10^3\rightarrow10^4$ decreases the RMS displacement in 10~yrs due to the unmodelled MBAs from $\sigma_\phi=1300\rightarrow60\rightarrow10$~m, with the radial and vertical RMS being substantially lower.  These correspond to 0.7, 0.03, and 0.005~mas, respectively.  Incorporating $N=1000$ MBAs into the ephemeris model would render noise from the rest of the asteroids unimportant for nearly all uses of Gaia or LSST MBA data for inference.  For occultation-based positions, one would probably want to have the $\approx3000$ most massive MBAs as active bodies with masses as free parameters in the ephemeris model in order to render encounters from unmodelled bodies as unimportant.

\subsection{Comparison to radiation forces}
As we consider smaller MBAs as test particles, radiation pressure becomes larger relative to gravitational forces.  It is interesting to compare the effects of radiation pressure to those of Brownian motion to see when the former are a bigger source of noise than the latter, or contrarily to see whether Brownian motion interferes with attempts to measure radiation forces on MBAs.  At first order, radiation pressure from the Sun and the reaction from reflected sunlight cause a constant radial force on the MBA.  The ratio of this force to the gravitational force for an asteroid of radius $R,$ geometric albedo $q$, and density $\rho$ is $(1+q)\Psi,$ where
\begin{equation}
 \Psi \equiv \frac{F_{\rm rad}}{F_{\rm grav}} = \frac{\pi R^2 L_\odot/4\pi c}{4\pi GM_\odot R^3\rho/3}
   = \frac{3}{16\pi} \frac{L_\odot}{GM_\odot c R \rho} \approx 3\times10^{-10} 10^{0.2(H-16)},
\end{equation}
where the last expression adopts our standard values $q=0.25, \rho=2500\,{\rm kg}\, {\rm m}^{-3}.$ The observational consequence of this constant radial force is an $H$-dependent shift $\Delta r = -(1+q)\Psi a$ at a fixed period of rotation, which does not resemble any expected effect of Brownian motion.  More important is the Yarkovsky effect, in which anisotropic re-emission of the absorbed radiation leads to a net force that is fraction $g_Y(1-q)$ of the incident radiation pressure.  The Yarkovsky effect has been detected for several NEAs but no individual MBAs, with values $g_Y \approx 0.1$ \citep[Yarkovsky effect and data are review by][]{yarkovsky}.  This is also consistent with detections of the Yarkovsky effect by examining the $\Delta a$ dependence on $H$ of members of asteroid \emph{families} in the Main Belt \citep[\eg][]{karin}.

 The Yarkovsky effect has in common with Brownian noise that it has the strongest observable effects via changes to $a$ and the mean-motion rates, and that these changes can be of varying sign and amplitude for a given asteroid.  In the case of the Yarkovsky effect, this variability arises because of the random orientations of spin axes, and differences in spin rates, shapes, and surface properties.

Taking a case where the Yarkovsky force to be in the prograde azimuthal direction, and using our system of units with $a=1$ and $v_c=1,$ this leads to a continuous power $dE/dt\approx g_Y\Psi$ which in turn implies $da/dt\approx2g_Y\Psi.$   The mean motion rate is hence $1-3 g_Y\Psi t$ and the accumulated shift in mean anomaly over time $T$ amounts to $\Delta\phi = -3 g_Y \Psi  T^2/2.$   In this case, a rough comparison of the accumulated $\Delta\phi$ from the Yarkovsky effect to the $\sigma_\phi$ of noise from unmodelled Brownian motion, after a time interval $T$, is
\begin{equation}
   \frac{\Delta\phi}{\sigma_\phi} \approx \frac{ 3 g_Y \Psi T^2/2}{\sqrt{3 T^3 \nIphi}} \approx \sqrt{T/10\,{\rm yr}} 10^{0.2(H-20.2)}
     \label{eq:yarkovsky}
\end{equation}
if we assume $\nIphi$ from the current knowledge of asteroid masses.   There is thus a break point near $H=20$, or $\approx100$~m diameter: for MBAs larger than this, the ephemeris uncertainties from the Yarkovsky effect will be smaller than those from mutual gravitational encounters on decade timescales. This includes all MBAs detected by Gaia and nearly all detectable by LSST.  Below this size, the Yarkovsky displacement grows larger in a decade and has a better chance of being detectable in the presence of Brownian motion.  Distinguishing the steady quadratic growth of the Yarkovsky $\Delta\phi$ from the random-walk growth of Brownian motion to the same final size requires observational S/N and time resolution significantly better than is needed for detecting either one individually.

When Yarkovsky forces on MBAs are estimated using the $H$-dependent dispersion of $a$ for asteroid families, the effective ``observation'' time period is typically several Myr.  Over these long time scales, the temporally coherent Yarkovsky effect can dominate over Brownian motion for much larger bodies, which is why the technique is successful.

@@

In conclusion, the Brownian motion from unmodelled or mis-modelled mutual impulses between MBAs causes RMS azimuthal displacements of $\approx2$~km (or 1 mas in angular position) over 10 years, with current knowledge of the masses of large asteroids.  This is large enough to become a significant part of the error budget for many inferences from Gaia MBA astrometry.  The knowledge on large-MBA masses likely to be gained from LSST tracking of MBAs should reduce this by a factor of $\ge4,$ at which point the Brownian noise will be comparable to Gaia or LSST measurement uncertainties only for the brighter targets of each survey.  In all cases, the Brownian motion is $\approx7\times$ smaller in the radial direction and $\approx45\times$ in the vertical direction.  The RMS azimuthal displacement grows with time as $T^{3/2}$, the vertical RMS as $T^{1/2}$, and the radial RMS eventual switches from the latter to the former at some time driven by the orbit eccentricity.

Making inferences from MBA astrometry becomes more complicated when Brownian motion is a significant contributor to the error budget, but a good path forward is to do all model fitting by including the $\approx1000$ largest MBAs as active bodies in the ephemeris model and leaving their masses as free parameters.  This will constrain the masses of these bodies to the full extent that the data allow, and will propertly propagate any remaining uncertainty in those masses into the other inferences that one is trying to make.  The noise contributed by the remaining millions of unmodelled accuracy would be just $\approx60$~m RMS in the azimuthal direction (about 30~$\mu$as), and a few $\mu$as or less in the other axes, small enough to ignore entirely.  
Occultation-based astrometry is much more accurate (but has fewer possible measurements per tracer), and would be significantly impacted by Brownian motion until $\gtrsim1000$ MBAs have mass estimates at modest $S/N$ level, although the precise impact will depend on whether the signals one is seeking are parallel or orthogonal to the kind of variations caused by Brownian motion.  For objects $\gtrsim100$~m in diameter, uncertainties from mutual encounters are larger than uncertainties from the Yarkovsky effect over decade timescales, at present levels of knowledge of asteroid masses.
